Assessing impact of geography as confounder in causal relationships between couples

### 2021-09-15 analysis/geographical_impact.Rmd

- GWAS of first 40 PCs run using [bgenie](https://jmarchini.org/bgenie/).
- First, inverse normal rank transformation (INT) was applied within genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`) before performing GWAS (this was done before dividing into sex-specific subgroups).
- After INT, PCs were residualized for `age` and `sex`.
- Run in:
  - Entire sample (residualized for both age and sex).  
  - In sex-specific subgroups (residualized for only age). 
- This was run in the entire genetic, consenting, European UKB sample (not just couples).
- GWAS was **not** adjusted for PCs (as is typical in GWAS) because PCs are linearly independent of each other (although the sample used for PC calculation - by the UKB investigators - would be slightly different that our GWAS sample, so the PCs would not necessarily be completely linearly independent in this sample).

@@

PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
DT::datatable(PCs_corr, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)

>>>>

- Performed GWAS of first 40 PCs, run using [bgenie](https://jmarchini.org/bgenie/).
- Sample selection: genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`), in attempt to best replicate the GWAS used in other analyses from the Neale lab. 
- Data transformation:
    - First, inverse normal rank tansformation (INT) was applied within the selected (this was done before dividing into sex-specific subgroups).
    - Next, PCs were residualized for `age` and `sex`.
- GWAS was run in:
    - Entire sample (residualized for both age and sex).  
    - Sex-specific subgroups (residualized for only age). 
- GWAS was **not** adjusted for PCs (as is typical in GWAS) because PCs are linearly independent of each other (although the sample used for PC calculation - by the UKB investigators - would be slightly different that our GWAS sample, so the PCs would not necessarily be completely linearly independent in this sample).

@@

PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)
